### 2022-06-24 temp.rmd
more

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r scEiaD_data_overview, fig.height = 18, fig.width = 10, echo = F, message = F}
plot_grid(a_aa_b, c_d , e, ncol = 3, rel_widths = c(1,1.6, 2.2))
```

```{r, fig.height=11, fig.width=6}
ComplexHeatmap::draw(hm, padding = unit(c(2, 2, 32, 2), "mm"))
```

```{r, fig.width=3, fig.height=7}
hm_flipflop
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.