### 2022-07-07 tester.Rmd
tweak fig sizes

---
title: 'PLAE enables quick and easy searching across one million ocular transcriptomes'
author:
  - Vinay Swamy:
      institute: bg
  - Zachary Batz:
      institute: nnrl
  - David McGaughey:
      institute:
        - bg
      correspondence: "yes"
      email: mcgaugheyd@mail.nih.gov
institute:
  - bg: Bioinformatics Group, Ophthalmic Genetics & Visual Function Branch, National Eye Institute, National Institutes of Health
  - nnrl: Neurobiology, Neurodegeneration & Repair Laboratory, National Eye Institute, National Institutes of Health
  - mgog: Medical Genetics and Ophthalmic Genomics Unit, National Eye Institute, National Institutes of Health

date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: zotero_plaeApp.bib
csl: investigative-ophthalmology-and-visual-science.csl
abstract: "PURPOSE: To create a high performance reactive web application to query gene expression across cell type, species, study, and other factors. METHODS: We updated the content and structure of the underlying data (sc Eye in a Disk, scEiaD) and wrote the web application PLAE (https://plae.nei.nih.gov) to visualize and explore it. RESULTS: The new portal provides quick visualization of over a million individual cells from vertebrate eye-and body transcriptomes, XX cell types, XX ocular tissues, XX body tissues. As a test of the value of this unified pan-eye dataset, we show XX. CONCLUSION: The PLAE v0.91 web app serves the pan-ocular and body dataset, scEiaD. This offers the eye community a powerful and quick means to test hypotheses on gene and transcript expression across 54 body and 19 eye tissues. "

keywords: "RNA-seq, lens, cornea, retina, RPE, Snakemake, web, app"
---

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)
library(tidyverse)
library(citr)
library(cowplot)
library(ggrepel)
library(colorspace)
library(flextable)
library(captioner)
library(glue)
library(tictoc)
tic()
data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD//'
data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
#data_dir_vGiga <- '~/data/scEiaD_gigascience/'
#data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
# setup caption-ing
fig_cap <- captioner("Figure")
supFig_cap <- captioner("Supplemental Figure")
tab_cap <- captioner("Table")
supTab_cap <- captioner("Supplemental Table")

library(pool)
library(RSQLite)
meta_filter_v0 <- fst::read_fst(glue("{data_dir_vGiga}/paper_data/2021_03_17_meta_filter.fst")) %>%

  mutate(CellType_predict = case_when(CellType_predict == 'AC/HC_Precurs' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType_predict),
         CellType= case_when(CellType == 'AC/HC_Precurs' ~ 'AC/HC Precursor',
                             TRUE ~ CellType))

meta_filter <- fst::read_fst(glue("{data_dir_vPLAE}/meta_filter.fst")) %>% filter(study_accession != 'Bharti_Nguyen_iRPE_2D_3D') %>%

  mutate(CellType_predict = case_when(CellType_predict == 'AC/HC_Precursor' ~ 'AC/HC Precursor',
                                      TRUE ~ CellType_predict),
         CellType= case_when(CellType == 'AC/HC_Precursor' ~ 'AC/HC Precursor',
                             TRUE ~ CellType),  
         Tissue = case_when(grepl('Iris', Tissue) ~ 'Iris',
                            TRUE ~ Tissue),
         Tissue = case_when(grepl('^RPE|^Choroid', Tissue) ~ 'RPE-Choroid',
                            TRUE ~ Tissue),
         Citation = case_when(study_accession == 'OGVFB_Hufnagel_iPSC_RPE' ~ 'Swamy VS, Fufa TD, Hufnagel RB, McGaughey DM Building the Mega Single Cell Transcriptome Ocular Meta-Atlas Gigascience 2021 Oct 13;10(10)',
                              TRUE ~ Citation),
         PMID = case_when(study_accession == 'OGVFB_Hufnagel_iPSC_RPE' ~ '34651173',
                          TRUE ~ PMID),
         study_accession = case_when(study_accession == 'OGVFB_Hufnagel_iPSC_RPE' ~ 'SRP329495',
                                     TRUE ~ study_accession))

scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vGiga}/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite'))
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vPLAE}/MOARTABLES__anthology_limmaFALSE___4000-counts-universe-study_accession-scANVIprojection-15-5-0.1-50-20__pointRelease01.sqlite'))

load('data/diff_data_pull.Rdata') # from src/diff_data_pull.R
toc()
```

```{r, fig.width=4, fig.height=15, echo = F, message = F}
# tic()
# # flip_flop diff genes
# down_tab <- down_tab_big %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(Base %in% ws_oct, 
#          Against == 'All')
# 
# org_diff_diff_genes <- diff_tab_big %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(gene_ct_combo %in% down_tab$gene_ct_combo, 
#          Against == 'All') %>% 
#   arrange(padj) %>% mutate(x = paste0(Symbol, ' (', ID, ')')) 
# 
# flip_flop_genes <- scEiaD_2020_v01 %>% tbl('diff_testing_genes') %>% collect() %>% filter(grepl(org_diff_diff_genes %>% pull(ID) %>% paste0(., collapse = '|'), Gene)) %>% pull(Gene)
# flip_flop_long <- scEiaD_2020_v01 %>%
#   tbl("diff_testing") %>%
#   filter(Gene %in% genes) %>%
#   collect() %>%
#   mutate(Symbol = str_extract(Gene, '^\\w+ ') %>% gsub(' $', '', .)) %>% 
#   mutate(gene_ct_combo = paste(Symbol, Base, sep = '_')) %>% 
#   filter(gene_ct_combo %in% org_diff_diff_genes$gene_ct_combo,
#          Group == 'CellType (Predict)',
#          Against == 'All') %>%
#   select(Gene, Organism, log2FoldChange, Base)
# 
# flip_flop_wide <- flip_flop_long %>% pivot_wider(values_from = log2FoldChange, names_from = Organism)
# flip_flop_organism <- colnames(flip_flop_wide)[c(3,4,5)]
# flip_flop_wide <- flip_flop_wide %>% data.frame()
# row.names(flip_flop_wide) <- paste0(flip_flop_wide$Gene,'_', flip_flop_wide$Base)
# 
# flip_flop_wide <- flip_flop_wide[,-c(1,2)]
# 
# 
# 
# flip_flop_genes <- row.names(flip_flop_wide) %>% gsub(' \\(.*','',.)
# 
# flip_flop_base <- row.names(flip_flop_wide) %>% gsub('.*_','',.)

hm_flipflop <- ComplexHeatmap::Heatmap(flip_flop_wide, column_labels = flip_flop_organism, 
                                       row_split = flip_flop_base, show_row_dend = FALSE, 
                                       row_labels = flip_flop_genes, 
                                       column_title_rot = 90, 
                                       name = 'log2FC', use_raster = TRUE,
                                       row_title_rot = 0)
hm_flipflop
#toc()
```